---
title: "osu! Beatmap Stats"
output: 
  html_document:
    toc:  true
---

# Introduction

Basic data visualization of beatmap info provided by osu! API. https://github.com/ppy/osu-api/wiki

Only uses ranked/loved/qualified maps. Graphs focusing on standard mode. 

(A few maps, due to ranked/loved irregularities, are actually in graveyard but included in the API query anyway. Ex. https://osu.ppy.sh/b/766190&m=2 with 1 loved CtB diff. )

```{r}
library(ggplot2)
library(gridExtra)
library(plyr)
library(jsonlite)
library(varhandle)
library(chron)
library(magrittr)
```

```{r, cache=TRUE}
# Cached chunk
# Import data from JSON and remove duplicate rows
beatmaps <- unique(do.call("rbind", fromJSON("maps.json")))
```

```{r}
# Set global figure width and height
knitr::opts_chunk$set(fig.width=10, fig.height=6) 
# Disable warnings and messages
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# Convert strings of ints and floats to numeric datatypes
beatmaps.isintcol <- sapply(beatmaps, function(col) all(check.numeric(col, only.integer=TRUE))) 
beatmaps.isnumcol <- sapply(beatmaps, function(col) all(check.numeric(col))) & !beatmaps.isintcol
beatmaps[, beatmaps.isintcol] = sapply(beatmaps[, beatmaps.isintcol], as.integer)
beatmaps[, beatmaps.isnumcol] = sapply(beatmaps[, beatmaps.isnumcol], as.numeric)

# Convert MySQL datetimes to R datetimes. As far as the new site goes, this appears to be UTC-4, but
# this may be a local thing. Probably broken. 
beatmaps$approved_date <- as.POSIXct(beatmaps$approved_date, tz="Etc/GMT+4")
beatmaps$last_update <- as.POSIXct(beatmaps$last_update, tz="Etc/GMT+4")


# Create labels and data frames for each gamemode
gamemodes <- c("std", "taiko", "ctb", "mania")
gamemode.labels <- c("Standard", "Taiko", "CtB", "Mania")
beatmaps$mode <- factor(beatmaps$mode, labels=gamemode.labels)
for (i in 1:4) {
  assign(gamemodes[i], beatmaps[beatmaps$mode == gamemode.labels[i],])
}

# Various plot parameters for convenience
# These will usually leave a few outlier maps out
AR_y_scale          <- scale_y_continuous(breaks=seq(0,10))
SR_y_scale          <- scale_y_continuous(limits=c(0,10), breaks=seq(0,10,1))
approved_x_scale    <- scale_x_datetime(date_breaks="1 year", date_labels="%Y")

legend_title_fill   <- labs(fill="Mode")
legend_title_color  <- labs(color="Mode")

# Center titles
theme_update(plot.title = element_text(hjust = 0.5))
```

# Plots

## Star Rating

```{r}
diff_x_scale <- scale_x_continuous(limits=c(0,10), breaks=seq(0,10,0.5))
diff_hist <- geom_histogram(binwidth=0.05)

# Histogram of star rating (all modes)
ggplot(beatmaps, aes(difficultyrating, fill=as.factor(mode))) + 
  ggtitle("Total Star Rating (All Modes)") + legend_title_fill +
  diff_x_scale + diff_hist

# Frequency polygon of SR (all modes)
ggplot(beatmaps, aes(difficultyrating, color=as.factor(mode))) + 
  ggtitle("Star Rating (All Modes)") + legend_title_color +
  diff_x_scale + geom_freqpoly(binwidth=0.05) 

# Histograms of SR (all modes) with separate y scales
diffplots <- llply(1:4, function(m) {
  ggplot(beatmaps[as.numeric(beatmaps$mode) == m, ], aes(difficultyrating)) +
    diff_x_scale + diff_hist + 
    ggtitle(gamemode.labels[m])
})
grid.arrange(grobs = diffplots, top="Star Rating Distributions (All Modes)")
```

The ranking criteria until recently required non-marathon maps to have a difficulty Normal or below. 
According to the new criteria, any map with less than 3:30 drain time requires a Normal or below. 
This explains the large quantity of maps between 1\* and 2\*. 

After these Easy and Normal maps, the most popular standard maps are between 3\* and 4\*. There is a small but noticeable spike in standard maps at about 5.25\*.

Taiko stands out for having the most unimodal-looking distribution. 

## Total Length

```{r}
length_x_scale <- scale_x_continuous(limits=c(0,600), breaks=seq(0,600,30))
length_hist <- geom_histogram(binwidth=1)
x_labels_90 <- theme(axis.text.x = element_text(angle=90, hjust=1))

# Histogram of total length (all modes)
ggplot(beatmaps, aes(total_length, fill=as.factor(mode))) + 
  ggtitle("Total Beatmap Length (All Modes)") + legend_title_fill +
  length_x_scale + length_hist

# Frequency polygon of total length (all modes)
ggplot(beatmaps, aes(total_length, color=as.factor(mode))) + 
  ggtitle("Beatmap Length (All Modes)") + legend_title_color +
  length_x_scale + geom_freqpoly(binwidth=1)

# Histograms of total length (all modes)
lengthplots <- llply(1:4, function(m) {
  ggplot(beatmaps[as.numeric(beatmaps$mode) == m, ], aes(total_length)) +
    length_hist + length_x_scale + x_labels_90 + 
    ggtitle(gamemode.labels[m])
})

grid.arrange(grobs=lengthplots, top="Beatmap Length Distributions (All Modes)")
```

The massive spike in maps 85-90 seconds long corresponds to the ever-popular TV Size map genre.

This occurs in every gamemode, though mania and taiko have relatively large numbers of two minute maps. 

## Playcount

```{r}
# Frequency polygons of playcount (all modes)
ggplot(beatmaps, aes(playcount, color=as.factor(mode))) + 
  ggtitle("Playcount (All Modes)") + legend_title_color +
  scale_x_continuous(limits=c(0,1000000)) + 
  geom_freqpoly(binwidth=5000)
```

## Date approved

```{r}
beatmaps$month <- cut(beatmaps$approved_date, breaks="month")
year_x_scale <- scale_x_discrete(breaks=unique(cut(beatmaps$approved_date, breaks="year"))) 
x_labels_45 <- theme(axis.text.x = element_text(angle=45, hjust=1))

# Bar chart of date approved (all modes)
ggplot(beatmaps, aes(x=month, fill=mode)) + 
  ggtitle("Date Approved (All Modes)") +
  geom_bar(width=1) + 
  year_x_scale + x_labels_45

# Frequency polygon of date approved (all modes)
ggplot(beatmaps, aes(x=month, group=mode, color=mode)) +
  ggtitle("Date Approved (All Modes)") + 
  legend_title_color +
  geom_freqpoly(stat="count") +
  year_x_scale + x_labels_45
```

# Tables

## Most frequent artists, titles, sources, and creators

```{r}
library(knitr)
most.frequent.kable <- function(arr, lab) {
  # Neat use of pipe
  arr %>% table %>% sort(decreasing=TRUE) %>% head(20) %>% kable(col.names=c(lab, "Freq"))
}

beatmaps$artist  %>% most.frequent.kable("Artist")
beatmaps$title   %>% most.frequent.kable("Title")
beatmaps$source  %>% most.frequent.kable("Source")
beatmaps$creator %>% most.frequent.kable("Creator")
```

## Most favorited mapsets
```{r}
beatmaps %>% 
  subset(!duplicated(beatmaps$beatmapset_id)) %>%  # Keep rows with unique beatmapset_id
  arrange(desc(favourite_count)) %>% 
  head(50) %>% `[`(c("creator", "artist", "title", "favourite_count")) %>% kable
```


## Most played maps
```{r}
beatmaps %>%
  arrange(desc(playcount)) %>% 
  head(50) %>% `[`(c("creator", "artist", "title", "version", "playcount")) %>% kable
```

# Scatterplots

```{r}
# Scatterplot of AR vs BPM
ggplot(std, aes(bpm, diff_approach)) + 
  ggtitle("Approach Rate vs BPM") + 
  scale_x_continuous(limits=c(0,500)) + 
  AR_y_scale + 
  geom_point(alpha=0.1)

# Scatterplot of SR vs total length time
ggplot(std, aes(total_length, difficultyrating)) + 
  ggtitle("Star Rating vs Total Length") +
  length_x_scale + 
  SR_y_scale +
  geom_point(alpha=0.1)

# Scatterplot of max combo vs drain time
ggplot(std, aes(hit_length, max_combo)) + 
  ggtitle("Max Combo vs Drain Time") + 
  length_x_scale + 
  scale_y_continuous(limits=c(0,4000)) +
  geom_point(alpha=0.05)

# High linear correlation, as expected
summary(lm(max_combo ~ hit_length, data=std))

# Scatterplot of favorite count vs playcount
ggplot(std, aes(playcount, favourite_count)) + 
  ggtitle("Favorite Count vs Playcount") + 
  scale_x_continuous(limits=c(0,1000000)) +
  scale_y_continuous(limits=c(0,1000)) +
  geom_point(alpha=0.05)

# Scatterplot of playcount vs total length
ggplot(std, aes(total_length, playcount)) +
  ggtitle("Playcount vs Total Length") + 
  length_x_scale + 
  scale_y_continuous(limits=c(0,1000000)) +
  geom_point(alpha=0.1)

# Scatterplot of AR vs date approved
ggplot(std, aes(approved_date, diff_approach)) + 
  ggtitle("Approach Rate vs Date Approved") + 
  AR_y_scale + 
  approved_x_scale + 
  geom_point(alpha=0.05)

# Scatterplot of SR vs date approved
ggplot(std, aes(approved_date, difficultyrating)) +
  ggtitle("Star Rating vs Date Approved") +
  SR_y_scale + 
  approved_x_scale + 
  geom_point(alpha=0.1)
```

# Spread info

```{r}
# Playcount by song time, categorized by spread icon
# https://osu.ppy.sh/help/wiki/Difficulties#star-rating Not sure about values between boundaries

spread.sr = c(0, 1.51, 2.26, 3.76, 5.26, 6.76)  
spread.names = c("Easy", "Normal", "Hard", "Insane", "Expert", "Expert+")
spread.colors = c("olivedrab3", "paleturquoise", "gold", "hotpink", "purple", "darkgray")

# Assign difficulty rating by spread ranges to spread names 
beatmaps$spread_name <- spread.names[cut(beatmaps$difficultyrating, spread.sr, right=FALSE, labels=FALSE)]
std <- beatmaps[beatmaps$mode == "Standard",]  # update std 

hitlength.bins = seq(0, 360, 30)
par(mfrow=c(2,3), mar=c(4,4,4,1), cex.main=2)
for (i in 1:length(spread.names)) {
  std.spread = std[std$spread_name == spread.names[i], ]
  playcount.bin.sum = sapply(split(std.spread, cut(std.spread$hit_length, hitlength.bins)),
                             function(df) sum(df$playcount))

  barplot(playcount.bin.sum, space=0, width=30, xlab="Hit length (s)", ylab="Playcount Total", main=spread.names[i],
          col=spread.colors[i], axisnames=FALSE)
  axis(1, at=hitlength.bins)
}

# Same but 150+ hitlength and stacked bars
hitlength.bins.150 = seq(150, 360, 30)
playcount.bin.mat = matrix(ncol=length(hitlength.bins.150)-1, nrow=length(spread.names))
colnames(playcount.bin.mat) = head(hitlength.bins.150, -1)
rownames(playcount.bin.mat) = spread.names

for (i in 1:nrow(playcount.bin.mat)) {
  std.spread = std[std$spread_name == spread.names[i], ]
  playcount.bin.mat[i,] = sapply(split(std.spread, cut(std.spread$hit_length, hitlength.bins.150)),
                                 function(df) sum(df$playcount))
}

dev.off()  # Reset par 
barplot(playcount.bin.mat, space=0, width=30, col=spread.colors, xlab="Hitlength (s)", ylab="Total Playcount",
        legend.text=spread.names, axisnames=FALSE, main="Total Playcount by Hitlength and Difficulty")
axis(1, at=hitlength.bins.150-hitlength.bins.150[1], labels=hitlength.bins.150)
```


